RNeXML: a package for reading and writing richly annotated phylogenetic, character, and trait data in R

### 2015-01-05 manuscripts/manuscript.Rmd
Moves last paragraph of intro to conclusions.
Also tightens up future directions.

`RNeXML` not only allows R's ecosystem to read and write data in
the NeXML format through an interface that is no more involved than
reading or writing data from other phylogenetic data formats, but also
carries immediate benefits for its users. For example, comparative
analysis R packages and users frequently add their own metadata
annotations to the phylogenies they work with, such as annotations
of species, stochastic character maps, trait values, model estimates
and parameter values. `RNeXML` affords R the capability to harness
machine-readable semantics and an extensible metadata schema to
capture, preserve, and share these and other kinds of information, all
through an API instead of having to understand in detail the schema
underlying the NeXML standard. To assist users in meeting the rising
bar for best practices in data sharing in phylogenetic research
[@Cranston_2014], `RNeXML` captures metadata information from the R
environment to the extent possible, and applies reasonable
defaults.

@@

# Future Directions

The goals for continued development of `RNeXML` revolve primarily
around three directions. One is better interoperability with other
existing phylogenetic data representations in R, such those found in
the `phylobase` package [@phylobase]. Another is better integration of
the rich metadata semantics found in ontologies defined in the Web
Ontology Language (OWL), including programmatic access to machine
reasoning with such metadata. Finally, and similar to many other R
packages, including those for comparative phylogenetics, in light of
the increasing size of phylogenomic and other datasets mechanisms may
eventually be needed that avoid the requirement to load an entire
file's content into memory.

>>>>

To make the capabilities of NeXML available to R users in
an easy-to-use form, and to lower the hurdles to adoption of the
standard, we present RNeXML, an R package that aims to provide easy
programmatic access to reading and writing NeXML documents, tailored
for the kinds of use-cases that will be common for users and
developers of the wealth of evolutionary analysis methods within the R
ecosystem.

@@

# Conclusions and future directions

`RNeXML` allows R's ecosystem to read and write data in the NeXML
format through an interface that is no more involved than reading or
writing data from other phylogenetic data formats. It also carries
immediate benefits for its users compared to other formats. For
example, comparative analysis R packages and users frequently add
their own metadata annotations to the phylogenies they work with, such
as annotations of species, stochastic character maps, trait values,
model estimates and parameter values. `RNeXML` affords R the
capability to harness machine-readable semantics and an extensible
metadata schema to capture, preserve, and share these and other kinds
of information, all through an API instead of having to understand in
detail the schema underlying the NeXML standard. To assist users in
meeting the rising bar for best practices in data sharing in
phylogenetic research [@Cranston_2014], `RNeXML` captures metadata
information from the R environment to the extent possible, and applies
reasonable defaults.

The goals for continued development of `RNeXML` revolve primarily
around better interoperability with other existing phylogenetic data
representations in R, such those found in the `phylobase` package
[@phylobase]; and better integration of the rich metadata semantics
found in ontologies defined in the Web Ontology Language (OWL),
including programmatic access to machine reasoning with such metadata.
